---
title: "FishSET R package functions"
output:
   rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FishSET R package functions} 
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}  
editor_options: 
  markdown: 
    wrap: 72
---

<!--  pdf_document: default -->

<!--  html_notebook: default-->

<!--  word_document: default-->

<!--  html_document:-->

<!--    df_print: paged-->

\#\#Getting started The FishSET R package includes functions that can be
used in the R console or that can be run through the FishSET app (a
guided user interface built u sing the Shiny R package). Calling
function through the R console allows for greater flexibility and ease
in troubleshooting but requires knowledge of R and the FishSET R
package. The FishSET app is meant to allow for users to use functions
and models with limited knowledge of R. The FishSET app is opened by
typing `run_fishset()` in the R console. This tutorial does not use the
FishSET app but walks through examples of FishSET R package functions in
the R console.

\#\#Load the package.

```{r eval=FALSE}
install.packages('devtools')
library(devtools)
install.packages("PATH/TO/Directory/Containing/FishSET", repos=NULL, type='source')
library(FishSET)
```

\#\#Upload data When the data is loaded into the FishSET database a
series of checks are performed. The output of these checks will be
printed to the R console. Check these messages to identify if any
corrections should be made. In the box below, we load both the main data
set and the port data. We will also need a spatial data set. We will
load that with the `read_dat` function. The `read_dat` function
recognizes most data formats, including spss, json, and stata, and reads
the data into the working environment. The functions does not save the
data into the FishSET database.

```{r eval=FALSE}
#Upload data
load_maindata(dat, over_write = TRUE, project = 'EXAMP') 
load_port(dat, port_name = port, over_write = TRUE, project = 'EXAMP')
spatdat <- read_dat('')
```

Data are now loaded into the FishSET database. The first line of code
below returns a list of all the data tables in the FishSET database. The
second line of code returns the first six lines of data from the data
table you just loaded into the database.

```{r  eval=FALSE}
tables_database()
head(table_view('projectMainDataTable'))
```

Additional codes for working within the FishSET database:

```{r eval=FALSE}
#Check whether a table exists in the database.
  table_exists()      
#Remove a table from the database.
 table_remove()     
table_fields()      View column names of data table.
```

\#\#Data Quality FishSET has built-in functions to check for NAs, NaNs,
outliers, unique observations, empty variables, and whether latitude and
longitude variables are in the correct format. The `data_verification`
function checks for these potential issues. All messages returned to R
console are also saved in the log file.

```{r  eval=FALSE}
  data_verification()
  # The following functions are contained in `data_verification` but can be run on their own too.
    outlier_check()
    nan_identify()
```

Functions also exist to view table or plot output of data quality
assessments. All plots and tables are automatically saved with project
name to the `output` folder in the FishSET R package.

```{r  eval=FALSE}
#Table output functions
  summary_stats()
  outlier_table() 
  
#Plots output functions
  outlier_plot()
```

Functions to correct data quality issues.

```{r  eval=FALSE}
  na_filter()
  nan_filter()

  outlier_remove()

  unique_filter()

  empty_vars_filter()

  degree()
```

Data tables can be filtered using the `filter_table` and `filter_dat`
functions. The filter statements are saved as a table in the FishSET
database and can be applied to any dataset using the `filter_dat`
function. No function has been written to remove variables from the
working data set. Users should use base R functions for this. Within the
FishSET app variables can be removed from the working data set in the
`Data Exploration` tab. Use the `add_vars` function to add variables
from the raw data set back into the working data set.

```{r  eval=FALSE}
#Filter
  filter_table()
  filter_dat()

#Add variables back into the working data set
  add_vars()
```

\#\#Data Exploration\
View the spatial (`map_plot`) and temporal (`temp_plot`) distribution of
data and the relationship between variables (`xy_plot`). Plots and
tables to the `output` folder. The `map_kernel`, `getis_ord_stats`, and
`moran_stats` functions provide measures of clustering, or hot spot
analysis. The `density_plot` creates a density or cumulative
distribution function plot of the selected variable.

```{r  eval=FALSE}

#Map
  map_plot()
#map kernel
  map_kernel()
#Getis Ord
  getis_ord_stats()
#Moran's I
  moran_stats()
#Temporal Plots
  temp_plot()
#x-y plots
  xy_plot(dat, project, var1, var2, regress = FALSE) 
  density_plot()

```

\#\#Fleet Analyses Fleet functions focus on functions that allow users
to assign vessels to fleet and view fleet characteristics (vessel count)
and fleet effort (trip length, weekly effort), and catch as a fleet
(species catch, by catch, weekly catch). All functions return output as
plots and/or tables that are output to the R console and saved to the
`output` folder.

```{r  eval=FALSE}
#Define and store fleet expressions 
  fleet_assign(dat, project, cond = NULL, fleet_val = NULL, table = NULL, save = TRUE)
#Number of unique vessels by time period
  vessel_count(dat, project, v, t, period = "month", group = NULL, 
                         year = NULL, position = "stack", output = c("table", "plot"))
#Total species catch by period
  species_catch(dat, project, species, date, period = "month_abv", fun = "sum", group = NULL, year = NULL, 
                          convert_to_tons = TRUE, value = c("count", "percent"), 
                          output = c("table", "plot"), position = "stack", format_tab = c("wide", "long"))
#Compare bycatch to other species caught
  bycatch(dat, project, cpue, catch = NULL , date, names = NULL, group = NULL, 
                    year = NULL, period = "year", value = c("count", "stc"), 
                    output = c("table", "plot"), format_tab = "wide")
#Aggregate species catch by unique vessel ID
  sum_catch("MainDataTable", "myProject", "catch", "species == 'cod' & catch > .5", val = "per", out = "logical")
#Catch total by week
  weekly_catch(dat, project, species, date, year = NULL, group = NULL, fun = "sum", position = "stack",
                         convert_to_tons = FALSE, value ="percent", output = c("plot", "table"))
#Average CPUE by week
  weekly_effort(dat, project, cpue, date, group = NULL, year = NULL, plot_type = "line_point", 
                          output = c("plot", "table"), format_tab = "wide")
#Create a plot or table of vessel trip length
  trip_length(dat, project, start, end, units = "days", catch = NULL, 
                        hauls = NULL, output = c("table", "plot"), haul_to_trip = FALSE)
```

\#\#Simple Analyses\
View and quantify relationship between variables. The correlation
function, `corr_out` returns a plot and table of correlations values
between all numberic variables supplied. The regression function,
`xy_plot', returns a plot with fitted simple linear regression line and test statitistics whne regress is set to TRUE. When`regress\`
is set to FALSE, points are plotted without a best fit line.

```{r  eval=FALSE}
#correlation
  corr_out(dat, project, variables='all')
#regression
  xy_plot(dat, project, var1, var2, regress = TRUE)
```

\#\#Compute New Variables\
Modify or create variables such as cpue and trip centroid. Functions
focus on data transformations, generating ID variables (for haul or
fishery season), computing simple arithmetic functions such as catch per
unit effort or sum of hauls, generate dummy variablesm, create a variety
of spatial variables such as distance between hauls or the latitude and
longitude of the haul midpoint, and create trip level functions such
collapsing haul to trip and identfiying trip centroid.

```{r  eval=FALSE}
#Data Transformations
  ##Change time unit
  temporal_mod
  ##Coded variable based on quantiles
  set_quants()
#Nominal ID
  #Haul or Trip ID
  ID_var()
  #Fishery Seasonal Identifier
  create_seasonal_ID()
#Arithmetic and temporal
  #Numeric functions
  create_var_num()
  #CPUE
  cpue()
#Dummy
  #From variable
  dummy_num()
  dummy_matrix()
#spatial
  #Distance between two points
  create_dist_between()
  #Haul midpoint 
  create_mid_haul()
  #Duration of time
  create_duration()
  #Zone when choice where to go next was made
  create_startingloc()
#Trip-level Functions
  #Collapse Haul to trip
  haul_to_trip()
  #Trip distance
  create_trip_distance()
  #Trip centroid
  create_trip_centroid()
```

\#\#Zonal Definition\
Functions in this group require a spatial data frame defining fishing or
regulatory zone. Both the `find_centroid` and `assignment_column` can be
used on their own but are called as necessary in other function. The
`find_centroid` function identifies the centroid of a zone and the
`assignment_column` function assigns observations in the primary data
set to zones. The `create_alternative_choice` function defines
alternative fishing choices and is required.

```{r  eval=FALSE}
  find_centroid()
  assignment_column()
  create_alternative_choice()

```

\#\#Expected Catch/Revenue\
Compute expected catch or expected revenue. The function creates the
expected catch/revenue matrix based on user-provided parameters, and
three default cases: Near-term: Movng window size of two days. In this
case, vessels are grouped based on defineGroup parameter. Medium-term:
Moving window size of seven days. In this case, there is no grouping and
catch for entire fleet is used. Long-term: Moving window size of seven
days from the previous year. In this case, there is no grouping and
catch for entire fleet is used. Output is saved in the FishSET database
and called in the `make_model_design` function.

The `sparstable` function provides a table of data sparsity by time
periods. It should be used to assess the availability of data for
defining the temporal moving window size in the `create_expectations`
function. Sparse data is not suited for shorter moving window sizes.

```{r  eval=FALSE}
  sparsetable()  
  create_expectations()
```

\#\#Model design file and run The `make_model_design` function creates a
list with data and parameters required to run the models, which are
called in the `discretefish_subroutine` function. The
`discretefish_subroutine` requires inputs of initial parameters,
optimization options (maximum iterations, the relative tolerance of x,
report frequency, and whether to report iterations), the optimization
method chosen, which must be one of the base R `optim` options, and a
name for the model to allow for easier tracking and comparison of
models. The subroutine also requires `catch`, `choice`, and `distance`
data, plus any other data `otherdat` needed to run the chosen
likelihood, such as expected catches or harvester data. These data are
defined in the `make_model_design` file. Expected catches is pulled by
the `make_model_design` function from the saved table generated by the
`create_expectations` function. `make_model_design` calls output from
`create_alternative_choice` to generate distance data.

```{r  eval=FALSE}
make_model_design()
discretefish_subroutine()
```

\#\#\#Likelihood functions FishSET includes six functions. User-created
functions can be saved and logged using the `log_func-model` function.
`logit_c` Conditional logit likelihood `logit_avgcat` Average catch
multinomial logit procedure `logit_correction` Full information model
with Dahl's correction function `epm_normal` Expected profit model
normal catch function `epm_lognormal` Expected profit model lognormal
catch function `epm_weibull` Expected profit model Weibull catch
function

\#\#\#Model run output The discretefish_subroutine function can take 30
or minutes to run. Model run status is returned to the R console. The
subroutine function outputs model results in a list that can be
summarized as:

    errorExplain: If it exists, a description of the model error.
    OutLogit: A matrix of coefficients, standard errors, and t-statistics
    optoutput: Optimization information (such as number of function iterations)
    seoutmat2: Standard errors
    MCM: Model comparison metrics (e.g. AIC, BIC)
    H1: The inverse hessian

The output can be viewed using:

```{r  eval=FALSE}
#Display errorExplain output
globalcheck_view('pcodldglobalcheck20190604')
#Display all other model output.
model_out_view('pcodmodelout20190604')

```

\#\#\#Selecting models Model comparison metrics are conveniently saved
in a table in the FishSET database. To access the table use `model_fit`
with the project name. Identifying the best model for future reference
requires `select_model`. This function utilized R Shiny to open an
interactive window where best models can be selected and saved. Model
selection can also be accomplished from the discretefish_subroutine
function by setting the `select.model` parameter to TRUE. When the
parameter is set to TRUE, a interactive window opens after the model has
compiled and output has saved.

```{r  eval=FALSE}
model_fit('pollock')
select_model()
```

\#\#Interactivity R Shiny is used to provide interactive interfaces. All
FishSET functions described here can be run through a user interface
with `run_fishset_gui`. The user interace guides users through loading
data to model evaluation. In addition, users can use the `select_vars`
function to open an interactive interface for selecting a subset of
variables to keep in the working data set and the `add_vars` function to
interactively select vars to add back into the working data set from the
raw dataset.

```{r  eval=FALSE}
run_fishset_gui()
select_vars()
select_model()
add_vars()
```

## Reproducibility

FishSET was designed with the aim of reproducibility. All function calls
are logged in a dated file. Log files are stored in the `Logs` folder.
Each log call has a functionID and a list of parameters supplies (args).
Some logged functions includs kwargs, optional arguments, an output, or
a message. The message section is used to save text output from a
function call that users may want to reference later, such as the number
of number of rows with missing data.

For example, the function call

```{r eval=FALSE}
filter_table(dat = 'pcodmaindatatable', project = 'pcod', x = 'PERFORMANCE_Code', exp = 'PERFORMANCE_Code==1')
```

returns the following log entry:

```{r eval=FALSE}


 {
          "functionID": "filter_table",
          "args": [
            "pcodMainDataTable",
            "pcod",
            "PERFORMANCE_Code",
            "PERFORMANCE_Code==1"
          ],
          "kwargs": [],
          "output": "",
          "msg": [
            {
              "dataframe": "pcodMainDataTable",
              "vector": "PERFORMANCE_Code",
              "FilterFunction": "PERFORMANCE_Code==1"
            }
          ]
        }
```

Log entries are written in JSON. Future version of FishSET will include
a function that will read the log files and rerun function calls with
current or updated data.

Logging is built into FishSET functions. However, it is possible to
start a new log file using `log_reset`. New log files are started each
day. User-created functions, such as likelihoods, can be saved for
future use and logged using `log_func_model`.

\#\#\#Logging

```{r  eval=FALSE}
`log_reset`
`log_func_model`
```

In the FishSET app, each tab has a `Notes` sections that is saved to the
`Output` folder. Comments, observations, or thoughts typed by the user
in this section are saved in dated text files. Additional pre-build
messages are also saved, such as output from data evaluation functions.

Finally, the FishSET package includes a Notebook template for report
writing.
