---
title: "Create and Integrate Your Custom Model"
author: "FishSET Development Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Create your own likelihood function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The FishSET R package provides a robust framework for discrete choice
modeling (DCM), utilizing `RTMB` (R Template Model Builder) for
automatic differentiation and optimized parameter estimation. While
functions like `fishset_design()` and `fishset_fit()` offer a solid
foundation for conditional logit and expected profit models, we
recognize that specialized research often requires custom model
structures.

To maximize computational performance and ensure compatibility with
downstream FishSET tools (like our policy function and database
loggers), custom models must be integrated directly into the `RTMB`
framework. We encourage users to develop their own objective functions
and submit them to be merged into the official FishSET toolkit.

This vignette outlines the workflow for developing a custom model by
leveraging existing FishSET data pipelines and writing an `RTMB`
negative log-likelihood function. Reach out to our team at
[nmfs.fishset\@noaa.gov](mailto:nmfs@fishset.gov) with any questions.

## Development guidelines

1.  Fork the [FishSET repo](https://github.com/noaa-nwfsc/FishSET) and
    clone the forked repository to your computer. More information on
    "forking" repositories and the GitHub workflow can be found
    [here](https://git-scm.com/book/en/v2/GitHub-Contributing-to-a-Project).

2.  Open the FishSET_RPackage.Rproj file.

3.  Develop your custom data processing and `RTMB` objective function
    following the steps below.

4.  Modify `fishset_fit.R` and `fishset_design.R` locally to include
    your model.

5.  Review and test code on local datasets.

6.  Commit, push changes, and submit a [pull
    request](https://docs.github.com/en/get-started/exploring-projects-on-github/contributing-to-a-project#making-a-pull-request).

## Step 1: Extract Formatted Data

Rather than formatting data manually, rely on the built-in
`format_model_data()` function to aggregate your main dataset, distance
matrices, expectations, and gridded data.

This function saves the combined dataset as a `.qs2` or `.rds` file in
the `Models/FormattedData` directory inside your project folder. Load
this dataset into your environment to begin designing your model matrix.

```{r function1, eval = FALSE}
# Example: Loading formatted data
project_dir <- file.path(locproject(), "MyProject")
formatted_dir <- file.path(project_dir, "Models", "FormattedData")
my_data <- qs2::qs_read(file.path(formatted_dir, "MyProjectLongFormatData.qs2"))
```

## Step 2: Mock the Design Object

The `fishset_fit()` function expects a specific list structure, of class
`fishset_design`, to perform estimations. If your model requires
predictors or matrix shapes that the standard `fishset_design()`
function cannot currently generate, write a custom data processing
script that builds a list mirroring this required structure.

Your design object must contain the following elements:

-   `y`: A numeric vector indicating the binary choice variable (0 or
    1).

-   `X`: A sparse model matrix containing your formatted predictors.

-   `settings`: A list containing dimensions such as `N_obs` (number of
    observations), `J_alts` (number of alternatives), and `K_vars`
    (number of parameters).

-   `ids`: A list containing unique observation IDs and zone identifiers
    to be used for post-estimation predictions.

## Step 3: Write the RTMB Objective Function

The core of FishSET's speed relies on `RTMB::MakeADFun` and
`stats::nlminb`. To integrate your model, you must write a negative
log-likelihood (NLL) function compatible with this framework.

Inside `fishset_fit()`, parameter estimation relies on two main
components passed to `RTMB`:

-   `data_list`: A list containing all known data (`X` matrix,
    observations, dimensions).

-   `start_pars`: A list of initial parameter values to be optimized
    (e.g., `betas`).

Your custom function must extract these elements using `RTMB::getAll()`
and calculate the NLL.

### Template: Custom Objective Function

```{r function2, eval = FALSE}
# Example custom model template for an RTMB NLL function
nll_func_custom <- function(pars) {
  # Unpack data_list and parameters into the function environment
  RTMB::getAll(data_list, pars)
  
  # Perform matrix multiplication (e.g., sparse design matrix %*% parameters)
  v <- X %*% betas
  
  # Extract utility of chosen alternatives
  v_chosen <- v[chosen_lin_idx]
  
  # Reshape utility vector for log-sum-exp calculations
  dim(v) <- c(J_alts, N_obs)
  
  # Calculate log-sum-exp (consider robust numeric scaling if necessary)
  log_sum_exp <- log(RTMB::colSums(exp(v)))
  
  # Calculate the negative log-likelihood
  nll <- -sum(v_chosen - log_sum_exp)
  
  return(nll)
}
```

**Important Note**: When writing an `RTMB` function, you must use `RTMB`
compatible math functions (e.g., `RTMB::dnorm`, `RTMB::rowSums`) rather
than standard base R variants to ensure automatic differentiation works
properly.

## Step 4: Integrate and Submit

Once you have a working design object and an `RTMB` objective function,
you can integrate your code into the package.

1.  **Modify** `fishset_fit.R`: Add a new `if/else` block within the `fishset_fit()` function to handle your custom model type. Set up your unique `data_list` and `start_pars`, and define your custom `nll_func` within this block.

2.  **Standardize Outputs**: Ensure your code allows `RTMB::MakeADFun` and `stats::nlminb` to run, producing a result formatted as class `fishset_fit`. This guarantees your model outputs a coefficient table, diagnostics, AIC/BIC, and predicted probabilities that seamlessly integrate with the project database.

3.  **Submit a Pull Request**: Give the Pull Request an informative name and provide a brief description of the mathematical structure of your model. Our team will review the code and notify you of any requested changes prior to merging.