---
title: "Enter title here"
author: "Enter name here"
date: "`r Sys.Date()`"
output: html_document
params:
  project: "Enter project name"
  date: "Enter date here"
  printcode: no
---
<!-- YAML section (above) -->

<!-- This is setup section -->
```{r, setup, include=FALSE}
# The setup code chunk should be the first chunk in your report
knitr::opts_chunk$set(echo = FALSE)
library(pander)
library(FishSET)

locuser <- "Enter/path/to/folder/" # Enter a filepath to a local folder (optional)

# for dev purposes
loc <- "C:/Users/bryce.mcmanus/Work/FishSET_RPackage"  
loc2 <- paste0(loc, '/inst')
```

<!-- Delete text below before generating report -->
# Template Introduction

This is the FishSET R package report template. It is written in R Markdown and allows for output from developing and running location choice models using the FishSET R package to be easily incorporated into a shareable report. R Markdown is a powerful tool that allows users to incorporate text, code, and code output within a single document. Reproducibility is a core feature of R Markdown. It is also a useful tool for automatic report generation, which is covered in the “Reusing reports” section of this template.

The template is divided into sections, each with a series of guiding questions and suggestions for your analysis. These questions will prompt you to explain how important aspects of your analysis were conducted, such as how the data was prepared, how models were defined, and which plots and tables to include. This is a suggested report structure. Sections and subsections can be added, deleted, or moved. Text can be added or deleted from this template as needed. See "Generating reports" for more information.

The template is best used in conjunction with analyses. Values, plots, tables, and other output from the analysis can be directly inserted into the document. See "Formatting text" section below for details on formatting text and "Report functions" later in this document for details on inserting output from analyses and R code. When the report is done, generate (or “knit”) the document (see the "Generating notebook reports" section).

Guiding comments are provided between “<!-” and "-–>“. These comments will not show up in the output document. Use them to provide notes you want to save but do not want in the report. Guidance in R code is commented out with ”#".

Guidance on using the template to write a report is divided into report options, formatting text, working with R, inserting FishSET output into the report, and generating the report. This guidance text should be deleted before generating the report. After the guidance text we have included a suggested report outline with guiding questions and example code chunks.


## Report options and setup

The template has two separate sections that set the global options for the report: the YAML section and the setup code chunk.

The YAML section (example shown below) contains report metadata and appears at the very beginning of the template. It begins and ends with three dashes `---` and contains fields such as "Title", "Author", "Date", and "Output". Use quotation marks when entering the title, author, and date fields. The "Output" field determines the type of document that will be generated. By default, the output is set to `pdf_document` (see "Generating reports" for PDF requirements). To change the output, replace `pdf_document` with a valid document type, such as `word_document` or `html_document`. HTML output renders very quickly and is useful for previewing the report. Rendering a Word doc or HTML file can also be done by clicking the "Knit" menu on the report's toolbar and selecting the desired output.  

<!-- 
---
title: "FishSET report"
author: "Jane Doe"
date: "`r Sys.Date()`"
output: pdf_document
params:
  project: "pollock2015"
  date: "2020-10-16"
  printcode: no
---
-->

The YAML section also contains an optional field called `params` (short for parameters) which allows users to create a list of information that can be used throughout the document. The `params` field is not required; however, it can be very useful. Entering key project information once at the beginning of the report can reduce the number of typos and errors, as well as make the report easier to reuse in the future. The code chunk below demonstrates how these parameters can be accessed in the report: 

```{r}
params$project
params$date
```

Here is an example of using parameters in a report function:

<!--
```{r}
 pull_table(project = params$project, fun = "vessel_count", date = params$date)
```
-->

The setup code chunk appears below the YAML section and should always be the very first code chunk in the report. This chunk sets the default behavior of each code chunk in the report. For example, the report template sets `echo = FALSE` which means that code contained in a code chunk will not appear in the report, only its output will be displayed. The setup chunk is also a good place to include any R packages that were used in the report. The `FishSET` and `pander` packages are required and are included in the template.

<!--
```{r, setup, include=FALSE}
# The setup code chunk should be the first chunk in your report
knitr::opts_chunk$set(echo = FALSE)

library(pander)
library(FishSET)
```
-->

To change the default option of specific chunks, add the option to the code chunk header:

<!--
```{r echo=TRUE}
Code
```
# Output
-->

In this example, the code will be shown above the output when the report is rendered. See https://bookdown.org/yihui/rmarkdown/r-code.html for more details. 

Additional resources:
Adjusting figures       [link](http://zevross.com/blog/2017/06/19/tips-and-tricks-for-working-with-images-and-figures-in-r-markdown-documents/)


## Formatting text

This section provides some basic tips for creating the FishSET reports with R Markdown. 

R Markdown documents are composed of text and code chunks. Text is written directly into the document. Formatting text is done inline using special characters and symbols.

Bold text         **text**
Italicized text   *text*
Bold and italic text	***text***
Block quote       > text
Endash            --
Emdash            ---
Subscript         x^~2~
Superscript       x^2
Equation          $A =\pi*r^[2]$
Equation block          <!-- $$E = mc^[2]$$ -->
Bulleted lists    * item one
                  * item two
                      + sub-item 1
                      - sub-sub item 1
Ordered list 	1. Item one
		          2. Item two
			            i) sub-item
                      A. sub-sub item
footnote          text[^1]
                  [^1]: Footnote text
Link  		[link](www.rstudio.com)                 
Paragraph break   Double-space at the end of a paragraph. <br>: each <br> corresponds to one line. 

See [link](https://bookdown.org/yihui/rmarkdown/markdown-syntax.html) or [link](https://rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf) for more information.   


## Integrating R Code

R code can be directly incorporated into the document. Doing so increases reproducibility and transparency while making updating the report much easier. Code chunks are used in this template to insert plots, tables, and other output into the report.  

To run and display R code in the report use inline formatting or code chunks. Inline code is identified using backticks, for example `code`. Code chunks are identified by two sets of three backticks, ```{r} code ```. To create a code chunk, you can type the code in the document, click the “Insert” drop down menu in the right-hand corner of the document window and select “R”, or use the keyboard shortcut “Ctrl + Alt + I” for Windows and “Command + Option + I” for Mac. This is an example of a R code chunk:

```{r}
2 + 2
```

To run the code before knitting the document, click on the green arrow in the far right-hand corner of the chunk. The output will be shown directly beneath the chunk. Code output can also be incorporated into the body of text by using inline R code: `r 2 + 2`. Code chunks can be easily adjusted or updated and ran as many times as necessary. This feature allows you to see code output without having to render the entire document, allowing you to quickly identify errors and potential issues such as typos or inserting the wrong plot. 

There are a number of code chunk options. Chunk options are written into the header and separated by commas. 
For example 
```{r, echo=FALSE, eval=FALSE}
2+2
````

General code chunk options:
`include = FALSE`     Prevents code and results from appearing in finished file. Use this when you don’t want the results displayed on their own but need the results in another code chunk. Default is TRUE - display results.
`eval = FALSE`    Prevents code from running. Default is TRUE - run the code chunk. 
`echo = FALSE`    Display results but not the code. This is generally preferred for tables and figures. Default is TRUE.
`message = FALSE`   Prevents messages generated by the code from appearing in the finished file. This is the default.
`warning = FALSE`   Prevents warnings generated by the code from appearing in the finished file. This is the default.

Use code chunk figure options to adjust the appearance of figures in the final document.
`fig.cap = “...”`   Add a caption to graphical results.
`fig.height = VALUE`	
`dev= ‘png’`    R function that will be used as a graphical device to record plot. Default is ‘png’.
`dpi = 72`    When applicable, dots per inch in graphics
`fig.align = ‘default’`   Align graphics to ‘left’, ‘right’, or  ‘center’.
fig.height    Adjust figure height (in inches)
fig.width   Adjust figure width (in inches)


## Report functions

The FishSET template utilizes three types of report functions to view and insert output created in FishSET: 

* Output functions
* Model summary functions
* Project summary functions

The output functions retrieve plots, tables, and notes created using FishSET. These functions begin with the prefix **pull** or **insert**. Functions with the prefix **pull** retrieve files from the package's output folder where plots and tables are automatically saved. Place **pull** and **insert** functions inside a code chunk to display output. Functions with the prefix **insert** will allow you to display output saved to a personal folder.

`pull_output` is the primary output function--it retrieves plots, tables, and notes from the FishSET Output folder given a project name, date, function name, and output type.

<!--
```{r}
pull_output(project = "pollock", fun = "density_plot", date = "2020-08-11", 
            type = "plot")
```
-->

`pull_plot`, `pull_table`, and `pull_notes` are the same as `pull_output` but each retrieves a specific output type.

<!--
```{r}
 pull_plot(project = "pollock", fun = "density_plot", date = "2020-08-11")
```
-->

In the above examples, outputs are the same: the plot for the function `density_plot` generated on 2020-08-11 for project "pollock".

There are two **insert** functions:  `insert_plot` and  `insert_table`.

Both **insert** functions require an object named `locuser` containing a valid folder directory exists in the global environment (for more information on finding a folder directory, see [insert helpful link here]). Enter the name of the file located in your folder and it will be displayed in the report:

<!--
```{r}
locuser <- "Enter/path/to/folder/"

# Enter the plot/table name from folder 
insert_plot("example_density_plot_gear.png")

insert_table("example_catch_by_gear.csv")
```
-->

The `locuser` object has been pre-written in the setup chunk (see "Report options and setup" below). Delete the # sign and enter a valid file directory inside the quotes. 

The Model summary functions display model output saved to the FishSET database. `model_out_summary`, `model_error_summary`, and `model_fit_summary` show model output, error, and fit tables given a project name. 

Project summary functions display the various database tables, files, and function calls that were created in the project. These are located in the "Appendix" section. They are pre-written and only require that the `project` and `date` fields have been entered in the YAML header (see "Report options and setup").


## Generating reports

R Markdown is capable of creating several types of documents, includeing PDF and Microsoft Word. Generating PDF documents requires a LaTeX engine which can be installed by running the code below:

```{r}
# install.packages("tinytex")
# tinytex::install_tinytex()  # install TinyTeX
```

The FishSET report template generates PDFs by default (see "Report setup" for changing output type). 

To generate your report, click on the “Knit” icon located on the toolbar or by using “Ctrl + Shift + K” (“Cmd + Shift + K” on macOS). Another option is to run rmarkdown::render() in the console. 

Additional resources:

generating MS Word documents    [link](https://bookdown.org/yihui/rmarkdown/word-document.html)
generating PDF documents        [link](https://bookdown.org/yihui/rmarkdown/pdf-document.html)
R Markdown Cheatsheet           [link](https://rstudio.com/resources/cheatsheets/)


<!-- Delete text above before generating report -->

# Introduction


## Scope of Analysis
<!-- 
* What are your main questions?

* What vessels, sectors, or fishers do you want to study? 

* What issues do you want to understand? Why are you conducting this analysis? 

* Which model(s) do you plan to use in the analysis? 
--> 


# Data


## Guiding Questions
<!-- 
* What data are needed? 

* Which variables do you need to explore your questions?

* Which model(s) do you plan to use? 
-->


## Source
<!-- 
* Note general source (e.g., AKFIN, AFSC, etc)

* Note price details (if applicable)
-->


## Characteristics
<!-- 
* Size (number of observations/individuals)

* Timeframe

* Spatial resolution and spatial data process

* Note the nature of integrated data and change
-->

<!-- summary table of data -->
```{r}
summary_table(params$project, output = "print")
```


## Data Exploration Description


### Fleet characteristics


### Vessels, years


### Maps / Spatial plots and tables


### Species plots 
<!-- if applicable -->


### Temporal plots and tables


# Methods


## Processing (data evaluation)


### Quality Checks

<!--
* Were NAs, NaNs or outliers found?
* Are all observations unique?
* Are any variables empty?
* Were any variable classes changed?
* Are latitude and longitude in correct format?
-->

<!-- 
* How were the data corrected?
  + Were any values removed or replaced?
    - If replaced, what replacement method was used?
-->

<!-- 
How were irregular data points handled (i.e. those occurring on land or outside zone)?
-->


## Data preparation
<!--
* Was the data filtered? 

* Were any variables jittered or randomized for confidentiality? 
-->


### CPUE calculation
<!-- 
* How was CPUE calculated? Which variables were used? Which unit(s) of time? 

* Include any other created or edited variables used in the analysis/models
-->


### Correlations


### Additional data 
<!-- 
* Was seasonal, auxiliary, or gridded data added to main data?
-->


## Models


### Alternative choice matrix


#### Parameters


### Expected catch/revenue matrix  
<!-- if applicable -->


#### Parameters


### Model background
<!-- 
* Were additional explanatory variables included, including interactions and gridded data?
* Which models were run with what data and parameters?
* Did models converge?
-->

<!-- record of model inputs -->
```{r}
function_summary(date = params$date, type = "model", show = "all")
```


```{r}
view_model_design(params$project, params$date)
```


# Results


## Hot Spot Plot (Spatial Kernel)


## Fleets
<!--
* How were fleets defined?
-->

```{r}
view_fleet_table(params$project)
```


### Fleet Analysis


### Fleet Comparison


## Model Results


### Parameters


### Values

```{r}
model_out_summary(params$project, output = "print")
```

### Errors 
```{r}
model_error_summary(params$project, output = "print")
```

## Model Comparisons

```{r}
model_fit_summary(params$project, output = "print")
```

# **Appendix A: Project Notes**

```{r}
pander(
  pandoc.list(
    pull_notes(params$project)
  )
)
```

<br>
<br>
<br>

# **Appendix B: Log Summary**

## Data Loading
```{r dat-load}
pander(
  function_summary(date = params$date, type = "dat_load", show = "last")
)
```

<br>
<br>

## Data Quality Evaluation
```{r dat-qual}
pander(
  function_summary(date = params$date, type = "dat_quality", show = "last")
)
```

<br>
<br>

## Data Exploration
```{r dat_explore}
pander(
  function_summary(date = params$date, type = "dat_exploration", show = "last")
)
```

<br>
<br>

## Fleet Assignment and Summary
```{r fleet}
pander(
  function_summary(date = params$date, type = "fleet", show = "last")
)
```

<br>
<br>

## Zonal Definition
```{r zonal-def}
pander(
  function_summary(date = params$date, type = "zonal_def", show = "last")
)
```

<br>
<br>

## Models
```{r model-fun}
pander(
  function_summary(date = params$date, type = "model", show = "last")
)
```

<br>
<br>
<br>

# **Appendix C: Project database tables**  

```{r}
project_tables(params$project)
```

<br>
<br>
<br>

# **Appendix D: Project file output (package output folder)**

```{r}
project_files(params$project)
```




