---
title: "Scallop End-to-End Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scallop End-to-End Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 72
---

## Overview

This vignette walks through the core FishSET workflow on the bundled
scallop example data. It starts with raw data loading, moves through QA/QC
and data reshaping, and ends with model fitting and diagnostics.

### Setup and data loading

We first load FishSET and the example datasets, then place the project in
a temporary working folder so the vignette can run without user-specific
paths. This keeps the example reproducible while still using the same
database-backed workflow as a real project.

```{r setup, include=FALSE}
library(FishSET)

folderpath <- tempdir()
project <- "scallop_vignette"

data("scallop")
data("scallop_ports")
data("tenMNSQR")
```

The example data cover three pieces of the workflow: trip-level records,
port locations, and the zone polygons used to assign fishing choices. A
quick look at their structure helps confirm the variables needed later in
the workflow are present.

```{r}
str(scallop[, c("TRIPID", "DATE_TRIP", "PERMIT.y", "DDLAT", "DDLON",
                "ZoneID", "LANDED_OBSCURED", "DOLLAR_OBSCURED")])
str(scallop_ports)
str(tenMNSQR, max.level = 1)
```

Now we load the raw data into FishSET using a single project name. The
functions save each table to the project database and make the working
table available in the R session for the next steps.

```{r}
load_maindata(dat = scallop, project = project, over_write = TRUE)
load_port(dat = scallop_ports, port_name = "port_name", project = project)
load_spatial(spat = tenMNSQR, project = project, name = "tenMNSQR")
```

## Quality assurance and quality control

We begin QA/QC by checking for missing values in the columns that matter
for modeling. Removing those rows up front ensures the later formatting
and model-fitting steps do not fail on incomplete observations.

```{r}
scallop_clean <- na_filter(
  dat = "scallop_vignetteMainDataTable",
  project = project,
  x = c("previous_port_lon", "previous_port_lat", "ZoneID"),
  remove = TRUE,
  over_write = TRUE
)
```

Next we run the spatial checks to identify observations that fall on land
or outside the intended zone domain. This also parses the date column so
we can summarize spatial issues over time. Note: removing observations
on land or outside the zone polygons is easier in the FishSET GUI.

```{r}
qaqc_out <- spatial_qaqc(
  dat = "scallop_vignetteMainDataTable",
  project = project,
  spat = "scallop_vignettetenMNSQRSpatTable",
  lon.dat = "DDLON",
  lat.dat = "DDLAT",
  date = "DATE_TRIP"
)

qaqc_out$spatial_summary
qaqc_out$land_plot
```

To see how the cleaned trips are distributed across zones, we summarize
the observations against the ten-minute-square grid. The tabular output
and static plot give a quick read on zone coverage before we reshape the
data for modeling.

```{r}
zone_out <- zone_summary(
  dat = "scallop_vignetteMainDataTable",
  spat = "scallop_vignettetenMNSQRSpatTable",
  project = project,
  zone.dat = "ZoneID",
  zone.spat = "TEN_ID",
  output = "tab_plot",
  plot_type = "static",
  dat_lon = "DDLON",
  dat_lat = "DDLAT"
)

zone_out$table
zone_out$plot
```

## Prepare and format data

The model needs a centroid table so it can measure distances from the
observed trip location to each alternative zone. Creating the zonal
centroid table also confirms that the spatial data and the trip zones
line up correctly.

```{r}
zone_centroid <- create_centroid(
  spat = "scallop_vignettetenMNSQRSpatTable",
  project = project,
  spatID = "TEN_ID",
  type = "zonal centroid",
  output = "centroid table"
)
```

We then define two alternative-choice sets with different minimum-haul
filters. Using two versions lets us compare how the available choice set
changes as we tighten the inclusion rules.

```{r}
create_alternative_choice(
  dat = "scallop_vignetteMainDataTable",
  project = project,
  alt_name = "alt1",
  zoneID = "ZoneID",
  occasion = "lon-lat",
  occasion_var = c("previous_port_lon", "previous_port_lat"),
  alt_var = "zonal centroid",
  min_haul = 5
)

create_alternative_choice(
  dat = "scallop_vignetteMainDataTable",
  project = project,
  alt_name = "alt2",
  zoneID = "ZoneID",
  occasion = "lon-lat",
  occasion_var = c("previous_port_lon", "previous_port_lat"),
  alt_var = "zonal centroid",
  min_haul = 200
)
```

Expected catch matrices translate the recent catch history into moving 
window averages for model inputs. We create one matrix for each 
alternative-choice definition so the later model specification can use 
the matching expectation set.

```{r}
create_expectations(
  dat = "scallop_vignetteMainDataTable",
  project = project,
  name = "exp_catch1",
  alt_name = "alt1",
  catch = "LANDED_OBSCURED",
  temp_var = "DATE_TRIP",
  temporal = "daily",
  temp_window = 7,
  day_lag = 1,
  year_lag = 0,
  empty_catch = NA,
  empty_expectation = 1e-4
)

create_expectations(
  dat = "scallop_vignetteMainDataTable",
  project = project,
  name = "exp_catch2",
  alt_name = "alt2",
  catch = "LANDED_OBSCURED",
  temp_var = "DATE_TRIP",
  temporal = "daily",
  temp_window = 7,
  day_lag = 1,
  year_lag = 0,
  empty_catch = NA,
  empty_expectation = 1e-4
)
```

The formatted data step reshapes the project tables into the long format
used by the RTMB model code within fishset_fit(). We keep the variables 
needed for the choice model and distance calculation, then repeat the 
process for the second alternative-choice set.

```{r}
format_model_data(
  project = project,
  name = "format_1",
  alt_name = "alt1",
  zone_id = "ZoneID",
  unique_obs_id = "TRIPID",
  select_vars = c("TRIPID", "DATE_TRIP", "ZoneID", "LANDED_OBSCURED"),
  expectations = "exp_catch1",
  distance = TRUE,
  distance_units = "mi"
)

format_model_data(
  project = project,
  name = "format_2",
  alt_name = "alt2",
  zone_id = "ZoneID",
  unique_obs_id = "TRIPID",
  select_vars = c("GEARCODE", "ZoneID", "DOLLAR_OBSCURED", "LANDED_OBSCURED"),
  expectations = "exp_catch2",
  distance = TRUE,
  distance_units = "mi"
)
```

## Model design and fit

The first design is a standard conditional logit with expected catch and
distance as fixed effects. The second adds area-specific constants so we
can compare a more flexible zonal logit specification.

```{r}
fishset_design(
  formula = chosen ~ exp_catch1 + distance,
  project = project,
  model_name = "clogit1",
  formatted_data_name = "format_1",
  unique_obs_id = "TRIPID",
  zone_id = "ZoneID",
  scale = TRUE
)

fishset_design(
  formula = chosen ~ exp_catch2 + distance + ZoneID,
  project = project,
  model_name = "zlogit1",
  formatted_data_name = "format_2",
  unique_obs_id = "TRIPID",
  zone_id = "ZoneID",
  scale = TRUE
)
```

The model-fitting step can take longer than a vignette render budget, so
we keep the fitting code in the document but do not evaluate it here. This
still shows the exact commands a user would run locally to estimate and
print both model tables. The printed tables include estimated coefficients,
standard errors, p-values, log-likelihood, AIC, Pseudo R2, and accuracy.

```{r eval=FALSE}
clogit1_fit <- fishset_fit(project = project, model_name = "clogit1")
print(clogit1_fit)

zlogit1_fit <- fishset_fit(project = project, model_name = "zlogit1")
print(zlogit1_fit)
```

## Model diagnostics and validation

Once the fits exist, we can test the IIA assumption with the
Hausman-McFadden test and then check whether residuals show spatial
correlation. These diagnostics help determine whether the final model is
well specified.

```{r eval=FALSE}
clogit1_iia <- fishset_iia_test(project = project, model_name = "clogit1")
print(clogit1_iia)

zlogit1_iia <- fishset_iia_test(project = project, model_name = "zlogit1")
print(zlogit1_iia)
```

```{r eval=FALSE}
clogit1_resid <- model_resid_corr(
  project = project,
  model_name = "clogit1",
  spat = "scallop_vignettetenMNSQRSpatTable",
  spat_id = "TEN_ID"
)
print(clogit1_resid)
plot(clogit1_resid)

zlogit1_resid <- model_resid_corr(
  project = project,
  model_name = "zlogit1",
  spat = "scallop_vignettetenMNSQRSpatTable",
  spat_id = "TEN_ID"
)
print(zlogit1_resid)
```
