---
title: "Scallop Conditional Logit Model Example"
author: "Bryce McManus"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Scallop Conditional Logit Model Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is an example of a conditional logit model using the `scallop` data
from the `FishSET` package.

### Packages

```{r}
library(FishSET)
```

```{r eval=TRUE, echo=FALSE}
# this chunk is for vignette version only
folderpath <- tempdir()
```


<br>
<br>

### Load Data

This analysis uses three of `FishSET`'s example datasets: the `scallop`
dataframe which contains anonymized scallop data from the Northeast,
`scallop_ports` which is a table of ports and their location, and
`tenMNSQR` which is a spatial dataframe of Northeastern ten minute
squares. 

The `scallop` dataframe contains a random sample of 10,000 trips 
from vessels in the Limited Access Days-at-Sea fleet when they 
are declared into either an Access Area or Open area Days-at-Sea
fishing trip.  

Noise was added to fishing locations, landing quantities, and the 
value of catch.  Permit, Operator, and trip identifiers were also anonymized. 



```{r}
data("scallop")
data('scallop_ports')
data("tenMNSQR")
scallop$LANDED_thousands<-scallop$LANDED_OBSCURED/1000
scallop$DOLLAR_2020_thousands<-scallop$DOLLAR_2020_OBSCURED/1000

vars_to_keep <- c('TRIPID', 'PERMIT.y', 'DATE_TRIP', 'DDLON', 'DDLAT', 'ZoneID', 
                  'LANDED_thousands', 'DOLLAR_2020_thousands', 'port_lon', 'port_lat',
                  'previous_port_lon', 'previous_port_lat')
scallop <- scallop[vars_to_keep]
```

<br>
<br>

Load each dataset into FishSET. Rescale the Landed weight to thousands of meat
pounds and the Dollar value to thousands of Real dollars. Note: when running 
this chunk you may get a pop-up asking to identify the location for a new 
FishSET folder or an existing folder.

```{r}
load_maindata(scallop, project = "scallopMod", over_write = TRUE)

load_port(dat = scallop_ports, port_name = "port_name", project = "scallopMod")

load_spatial(spat = tenMNSQR, name = "TenMnSqr", project = "scallopMod")

scallopModTenMnSqrSpatTable <- table_view('scallopModTenMnSqrSpatTable', "scallopMod")
```

<br>
<br>

### QAQC

`summary_stats()` is a useful way to find `NA`s in the data.

```{r}
summary_stats(scallopModMainDataTable, project = "scallopMod")
```

<br>
<br>

Remove `NA`s by calling `na_filter()`. In this case, three variables
which are important for modeling are being filtered: `ZoneID`,
`previous_port_lon`, and `previous_port_lat`.

```{r}
scallopModMainDataTable <- 
  na_filter(scallopModMainDataTable,
            project = "scallopMod", 
            x = c("ZoneID", "previous_port_lon", "previous_port_lat"),
            remove = TRUE)
```

<br>
<br>

Plot the number of observations by zone.

```{r}
zone_summary(dat = scallopModMainDataTable, 
             spat = scallopModTenMnSqrSpatTable, 
             project = "scallopMod",
             zone.dat = "ZoneID",
             zone.spat = "TEN_ID",
             output = "tab_plot")
```

<br>
<br>

Check for sparsity.

```{r}
sparsetable(scallopModMainDataTable, 'scallopMod', 
            timevar = 'DATE_TRIP', 
            zonevar = 'ZoneID',
            var = 'LANDED_thousands')

sparsplot('scallopMod')
```

<br>
<br>

### Create Centroids

A centroid table is needed to create the distance matrix. It can be used
as the choice occasion or the alternative choice. The simplest way to
create a zonal centroid table is by passing it to `create_centroid()`,
which saves the centroids to the FishSET database.

```{r}
create_centroid(spat = scallopModTenMnSqrSpatTable,
                project = "scallopMod",
                spatID = "TEN_ID",
                type = "zonal centroid",
                output = "centroid table")
```

<br>
<br>

### Alternative Choice

For this example, the alternative choice list will use the longitude and
latitude of the disembarking port (`previous_port_lon` and
`previous_port_lat`) as the choice occasion and the zonal centroid of
the fishing areas as the alternative. The minimum haul haul requirement
is set to 90.

```{r}
create_alternative_choice(dat = scallopModMainDataTable,
                          project = "scallopMod",
                          occasion = "lon-lat",
                          occasion_var = c("previous_port_lon", "previous_port_lat"),
                          alt_var = "zonal centroid",
                          zoneID = "ZoneID",
                          zone.cent.name = "scallopModZoneCentroid",
                          min.haul = 90
                          )
```

<br>
<br>

The plot below visualizes zone frequency after accounting for the
minimum haul requirement from the alternative choice list.

```{r}
z_ind <- which(alt_choice_list('scallopMod')$dataZoneTrue == 1)

zOut <- 
  zone_summary(scallopModMainDataTable[z_ind, ], 
               spat = scallopModTenMnSqrSpatTable, 
               project = "scallopMod",
               zone.dat = "ZoneID",
               zone.spat = "TEN_ID",
               output = "tab_plot")

pretty_tab(zOut$tab)

zOut$plot
```

<br>
<br>

### Expected Catch

This code chunk creates two different expected catch matrices: one using
a window of seven days, lag of one and a window of 14 days, lag of two. They
will be named `user1` and `user2` respectively.

```{r}
# user1 expected catch matrix
create_expectations(dat = scallopModMainDataTable,
                    project = "scallopMod",
                    name = "exp1",
                    catch = "LANDED_thousands", 
                    temp_var = "DATE_TRIP",
                    temp_window = 7,
                    day_lag = 1,
                    year_lag = 0,
                    temporal = 'daily',
                    empty_catch = NA,
                    empty_expectation = 1e-04)

# user2 expected catch matrix
create_expectations(dat = scallopModMainDataTable,
                    project = "scallopMod",
                    name = "exp2",
                    catch = "LANDED_thousands", 
                    temp_var = "DATE_TRIP",
                    temporal = "daily",
                    temp_window = 14,
                    day_lag = 2,
                    empty_catch = NA,
                    empty_expectation = 1e-04)
```

<br>
<br>

The data must be checked for common data quality issues before it can be
used in the modeling functions (i.e. `make_model_design()` and
`discretefish_subroutine()`). `check_model_data()` saves a new version
of the primary data with the suffix `_final` added to indicate that the
table is in its "final" state and ready to be used for modeling.

```{r}
check_model_data(scallopModMainDataTable, 
                 project = "scallopMod", 
                 uniqueID = "TRIPID",
                 latlon = c("DDLON","DDLAT"))
```

<br>
<br>

### Model Design

The model design file below will run two conditional logit models, each
using one of the expected catch matrices created earlier (this is
specified by using `'individual'` in the `expectcatchmodels` argument).

```{r}
make_model_design(project = "scallopMod",
                  catchID = "LANDED_thousands",
                  likelihood = "logit_c",
                  initparams = c(0, 0),
                  vars1 = NULL,
                  vars2 = NULL,
                  mod.name = 'lz', 
                  expectcatchmodels = list('individual')
                  )
```

<br>
<br>

### Run Models

Use `discretefish_subroutine()` to run all models in the model design
file.

```{r}
discretefish_subroutine(project = "scallopMod", explorestarts = FALSE)
```

<br>
<br>

Use `model_params()` to see the model output. `user1` and `user2` are the 
expected catch parameters and `V1` the travel distance parameter.  A reasonably
specified model should find positive coefficients for `user1` and `user2` and
negative coefficiencts for `V1`.

```{r}
model_params("scallopMod", output = 'print')
```

<br>
<br>

Compare model fit.

```{r}
model_fit_summary("scallopMod")
```

```{r eval=TRUE, echo=FALSE}
# for vignette version only
unlink(folderpath)
```
